Latent Variable Clustering for Classification of Events in the AT-TPC

### 2020-04-01 doc/Manuscript/article.tex
Updates from Overleaf
Started writing the introduction

\section{Introduction}\label{sec:intro}

\section{Methods}\label{sec:methods}
\subsection{Why machine learning}

>>>>

\section{Introduction}\label{sec:intro}
\begin{itemize}
    \item Overall goals for the AT-TPC
    \item Nuclear physics motivation for clustering-based (unsupervised) methods
    \item Motivation from ML approach

\end{itemize}
The AT-TPC \cite{Bradt2018} is a novel type of detector designed specifically for nuclear physics experiments where the energies of the recoiling particles are very low compared to the energy required to escape target material \cite{}. The luminosity of nuclear physics experiments performed with fixed targets is directly proportional to the amount of material encountered by the beam. On the other hand, for several classes of experiments the detection of recoil particles is paramount, therefore limiting the target thickness. In addition, the properties of the recoil particles are modified while traversing the target material, affecting the resolutions that can be achieved. This necessary balance between luminosity and resolution is particularly difficult when performing experiments with rare isotope beams, because of the low intensities available.

The concept of active target aims at mitigating this compromise, by turning the target itself into a detector \cite{Beceiro2018}. Most active target detectors such as the AT-TPC are composed of a time projection chamber (TPC) where the detector gas is at the same time the target material. Recoil particles that originate from a nuclear reaction between a beam nucleus and a gas nucleus can be tracked from the vertex of the reaction to their final position inside the active volume of the target. Their properties can therefore be measured without any loss of resolution regardless of the amount of material traversed by the beam. At the same time, the detection efficiency is dramatically increased by the very large solid angle covered in this geometry. A direct consequence of this concept is the inclusiveness of the experimental data recorded by this type of detector: any nuclear reaction happening within the target is recorded. Although this sounds like an advantage from the scientific point of view, it poses great challenges during the analysis phase, that are reminiscent of bubble chamber times and on par with nowadays event classification issues in particle physics. More often than not, the reaction channel of interest has one of the lowest cross sections, therefore the physicist is faced with the task of sorting out the corresponding events from the ``background" of other reaction channels.

Because TPCs produce 3-dimensional images of charged particle tracks, the event classification task is often akin to a visual inspection (bubble chamber times), which is not practical nowadays because of the large quantities of data. Machine learning techniques then appear as a promising prospect, in particular in the image recognition domain where much progress has been made recently. In addition, some ML algorithms offer new possibilities such as the potential discovery of unforeseen phenomena that would have been missed by more traditional analysis methods. The classification and clustering of events based on various ML algorithm is hereby examined, using experimental data recorded by the AT-TPC during its commissioning experiment from a radioactive $^{46}$Ar beam reacting of an isobutane target composed of proton and carbon nuclei.

\section{Experimental Details} 
(46Ar)
\section{Methods}\label{sec:methods}
\subsection{Why machine learning}
